---
title: "ice_cream_example"
output: html_document
---


#### NOTES
# do we need several samples to do this analysis? not just one sample... FIX

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 06

## Standard error of the mean

Calculate the standard error of the mean of an estimate of video score taken from 10 individuals. What does the standard error mean in words?

```{r}





```

## Confidence interval

Is it reasonable to calculate confidence intervals for this data? Why or why not?

```{r}



```

For your samples calculated above, calculate the confidence interval surrounding the estimation of the mean.

```{r}











```

# BONUS: Write a function to calculate the confidence interval above

```{r}

mean.conf.int <- function(x, CI = 0.95) {
    # define variables first
    
    
    
    
    # calculate interval now: xbar +/- t*std.error

        
    # return conf.int variable

}

# try out your function!


```

