---
title: "ice_cream_example"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 06

## Standard error of the mean

Calculate the standard error of the mean of an estimate of video score taken from 10 individuals. What does the standard error mean in words?

```{r}
set.seed(100)
# standard error = sd/sqrt(n)
video_sample <- sample(ice_cream$video, size = 10)

sd(video_sample)/sqrt(10)

print("The standard error of the mean is a measure of precision of the mean estimate from the sample.")

```

## Confidence interval

Is it reasonable to calculate confidence intervals for this data? Why or why not?

```{r}

print("The video data is approximately normally distributed and samples are random, independent, and taken from a much larger population. Therefore yes, it is appropriate to calculate the confidence interval.")

```

For your samples calculated above, calculate the confidence interval surrounding the estimation of the mean.

```{r}

# confidence interval (95%) = sample_mean +/- t_0.025 (sample_SE)

# first we need to find t_0.025, what are our degress of freedom?
# df = n - 1
df = 9

# can use the qt(p, df) function to find the test statistic
# p = 0.975 (1 - 0.025)
t <- qt(0.975, df)

# st error from previous question
se <- sd(video_sample)/sqrt(10)

# plug in!
mean(video_sample) - t*se
mean(video_sample) + t*se

# state conclusion:
print("We are 95% confident that the true mean in video scores lies between 45.78 and 62.61")

print(glue::glue("The actual mean of the video scores is: {mean(ice_cream$video)}, which is within our 95% confidence interval."))

```

# BONUS: Write a function to calculate the confidence interval above

```{r}

mean.conf.int <- function(x, CI = 0.95) {
    # define variables first
    xbar <- mean(x)
    n <- length(x)
    t.quantile <- qt(1-(1-CI)/2, df = n-1)
    std.error <- sd(x)/sqrt(n)
    
    # calculate interval now: xbar +/- t*std.error
    conf.int <- c(xbar-t.quantile*std.error, xbar+t.quantile*std.error)
    
    # return conf.int variable
    return(conf.int)
}

# try out your function!
mean.conf.int(video_sample)

```