---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 08

## Comparing populations - t test

Split the original population of individuals into males and females. Perform a statistical analysis to determine if males are more likely to score higher on the video than females.

```{r}

# split into two populations, male and female




# easy way - we can use t.test!




```

## Paired sample designs

Let's say we wanted to perform a statistical analysis to decide if the video game scores are statistically different from the puzzle scores. Would it be more appropriate to use an unpaired or paired-sample analysis? Why?

```{r}



```

Perform the analysis detailed above. What is the result?

```{r}

# lets calculate ourself first and then test with t.test
# first - calculate diff between video and puzzle scores



# now we calculate the mean and sd



# calculate the SE
# se = sd / sqrt(n) - n = 200, not 400!


# calculate confidence interval of mean estimate
# 95% CI = mean +/- t_0.025(SE)

# find t
# use qt(p, df)


# calculate the CI



# calculate test statistic t_s
# t_s = diff - 0 / SE


# calculate the p-value (area under the curve) at our test statistic - times by 2
# we want the area to the right of the t_s and to the left of -t_s, sometimes helpful to draw a quick sketch


# we can use t.test to verify we did it right... and we did! woo!


# our results:



```

## BONUS

Repeat the analysis above but with the other type of analysis (i.e. if you chose an unpaired test, repeat with a paired test). What do you notice that is different between the two?

```{r}

# lets try with t.test to save time...



```

## BONUS: Power

You would like to perform an observational study on this population of individuals to see if individuals who like vanilla scored higher on the video game than individuals who like strawberry ice cream (ignore chocolate ice cream fans for now). However, you are running out of time because your assignment is due tomorrow! Perform a power analysis to decide how many individuals you would need to sample in order to have high enough power (0.8) to detect a significant difference if one does exist. 

```{r}

# we can use power.t.test to calculate the power given n, diff, sd, and alpha
# first calculate power at one n: say n = 5
# we can estimate sd by using the population sd -- but for which population?
# we can set alpha = 0.05
# we can estimate diff by calculating the population diff





# calculate means of each group and then get the difference to estimate delta




# calculate power - choose one population to set the sd




# now we can calculate our n for a power of 0.8!



```

