---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 03

## Conditional probability

Using Bayes' theorem, find the probability that a randomly chosen individual is female given that their favorite flavor is 1) vanilla, 2) chocolate, or 3) strawberry

### Vanilla
```{r}

# bayes theorem
# P(A|B) = P(B|A)P(A)/P(B)

# get a table
table(ice_cream$ice_cream, ice_cream$female)

# vanilla:
# P(female | vanilla) = P(vanilla | female)P(female)/P(vanilla)
p_ab <- 48/(48+32+29)
p_a <- (48+32+29)/200
p_b <- (47+48)/200

p_ab*p_a/p_b

# check this with the table: p(female | vanilla)
48/(48+47)

```

### Chocolate
```{r}

# bayes theorem
# P(A|B) = P(B|A)P(A)/P(B)

# get a table
table(ice_cream$ice_cream, ice_cream$female)

# chocolate:
# P(female | choc) = P(choc | female)P(female)/P(choc)
p_ab <- 32/(48+32+29)
p_a <- (48+32+29)/200
p_b <- (15+32)/200

p_ab*p_a/p_b

# check this with the table: p(female | choc)
32/(32+15)

```

### Strawberry
```{r}

# bayes theorem
# P(A|B) = P(B|A)P(A)/P(B)

# get a table
table(ice_cream$ice_cream, ice_cream$female)

# strawberry:
# P(female | strawberry) = P(strawberry | female)P(female)/P(strawberry)
p_ab <- 29/(48+32+29)
p_a <- (48+32+29)/200
p_b <- (29+29)/200

p_ab*p_a/p_b

# check this with the table: p(female | strawberry)
29/(29+29)

```

## Random variables

What is the mean difference in video scores between males and females?

1. Find mean (and sd) for each population

```{r}

female <- ice_cream %>%
    dplyr::filter(female == 1)

male <- ice_cream %>%
    dplyr::filter(female == 0)

female_mean <- mean(female$video)
female_sd <- sd(female$video)

male_mean <- mean(male$video)
male_sd <- sd(male$video)

```

2. Find the difference in means

```{r}

# D = mu_1 - mu_2

female_mean - male_mean

```

3. Find the standard deviation of the difference in means between male and female video scores

```{r}

# first find difference in variance
# VAR = var_x + var_y
# VAR = sd_x^2 + sd_y^2

var_diff <- female_sd**2 + male_sd**2

# sd = sqrt(var)
sd_diff <- sqrt(var_diff)
sd_diff

```