---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 03

## Conditional probability

Using Bayes' theorem, find the probability that a randomly chosen individual is female given that their favorite flavor is 1) vanilla, 2) chocolate, or 3) strawberry

### Vanilla
```{r}

# bayes theorem


# get a table


# vanilla:



# check this with the table: p(female | vanilla)

```

### Chocolate
```{r}

# bayes theorem


# get a table



# chocolate:




# check this with the table: p(female | choc)


```

### Strawberry
```{r}

# bayes theorem


# get a table


# strawberry:




# check this with the table: p(female | strawberry)


```

## Random variables

What is the mean difference in video scores between males and females?

1. Find mean (and sd) for each population

```{r}

# filter only female data



# filter only male data



# find female mean and sd



# find male mean and sd



```

2. Find the difference in means

```{r}




```

3. Find the standard deviation of the difference in means between male and female video scores

```{r}

# first find difference in variance





# now find sd



```