---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 07

## Hypothesis testing

If we expect the sample mean to be equal to the population mean, what would be the null and alternative hypotheses in such an analysis?

```{r}

# create sample and se from last time
video_sample <- sample(ice_cream$video, 10)
se <- sd(video_sample)/sqrt(10)

# what are the hypotheses?


```

Find the test statistic, t, to choose between these two hypotheses above

```{r}






```

Given your sample, do you reject or fail to reject the null hypothesis?

```{r}






```

## Write a function for a one-sided t test

```{r}

one.sided.t.test <- function(x, mu = 0, lower.tail = TRUE) {
    # define variables first
    xbar <- mean(x)
    n <- length(x)
    std.error <- sd(x)/sqrt(n)

    # calculate t statistic

        
    # calculate pval

        
    # return pval variable

}

# try out your function!



```

## Type I and Type II Errors

Let's say your analysis leads you to reject the null hypothesis, but in reality the sample mean is equal to the population mean. What type of error have you made and how could you decrease the probability of this type of error?

```{r}






```

## Effect size 

Calculate the statistical effect size of the difference between the population and sample means. 

```{r}

# effect size = diff in means / sd




```
