---
title: "ice_cream_L01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

# load packages


```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

# read in the data
url <- "https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv"


# show data table

```

# Lecture 01

## What type of variables are in our dataframe?

* id: 
* female: 
* ice_cream: 
* video: 
* puzzle: 

## What is the shape, center, and spread of the four variables?

### Female

```{r}

# show table of counts


# plot gender



```

### Ice cream flavor

```{r}

# show table of counts



# plot ice cream flavor




```

### Video score

```{r}

# note we can use summary here instead of table because it is a continuous variable (not discrete)


# histogram





# measures of spread





# histogram with mean and sd plotted on top





# box plot







```

### Puzzle score

```{r}

# note we can use summary here instead of table because it is a continuous variable (not discrete)



# histogram





# redo histogram with less bins!





# measures of spread




# histogram with mean and sd plotted on top






# box plot






```