---
title: "ice_cream_L01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)

```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("~/Downloads/Ice_cream_R.csv")

DT::datatable(ice_cream)

```

# Lecture 01

## What type of variables are in our dataframe?

* id: numeric/discrete
* female: numeric/discrete
* ice_cream: numeric/discrete
* video: numeric/continuous
* puzzle: numeric/continuous

## What is the shape, center, and spread of the four variables?

### Female

```{r}

table(ice_cream$female)

ice_cream %>%
    dplyr::mutate(female = dplyr::recode(female, "0" = "Male", "1" = "Female")) %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = female, fill = female) +
    ggplot2::geom_bar(color = "black") +
    ggplot2::theme_bw(24) +
    ggplot2::scale_fill_manual(values = c("Female" = "pink", "Male" = "skyblue"),
                               name = "Gender") +
    ggplot2::labs(x = "Gender", y = "Count") +
    ggplot2::theme(legend.position = 'none')

```

### Ice cream flavor

```{r}

table(ice_cream$ice_cream)

ice_cream %>%
    dplyr::mutate(ice_cream = dplyr::recode(ice_cream, "1" = "Vanilla", 
                                            "2" = "Chocolate",
                                            "3" = "Strawberry")) %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = ice_cream, fill = ice_cream) +
    ggplot2::geom_bar(color = "black") +
    ggplot2::theme_bw(24) +
    ggplot2::scale_fill_manual(values = c("Vanilla" = "grey100", "Chocolate" = "chocolate4",
                                          "Strawberry" = "maroon3"),
                               name = "Gender") +
    ggplot2::labs(x = "Ice cream flavor", y = "Count") +
    ggplot2::theme(legend.position = 'none')

```

### Video score

```{r}

# note we can use summary here instead of table because it is a continuous variable (not discrete)
summary(ice_cream$video)

# histogram
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = video) +
    ggplot2::geom_histogram(bins = 20, fill = "chartreuse4") +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "Video score", y = "Count")

# measures of spread
var(ice_cream$video)
sd(ice_cream$video)

# histogram with mean and sd plotted on top
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = video) +
    ggplot2::geom_histogram(bins = 20, fill = "chartreuse4") +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "Video score", y = "Count") +
    ggplot2::geom_vline(xintercept = mean(ice_cream$video), color = "blue", size = 2) +
    ggplot2::geom_vline(xintercept = c(mean(ice_cream$video) + sd(ice_cream$video),
                                       mean(ice_cream$video) - sd(ice_cream$video)), color = "black", size = 1.5, linetype = 2)

# box plot
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = "1", y = video) +
    ggplot2::geom_jitter(width = 0.2) +
    ggplot2::geom_boxplot(alpha = 0.5) +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "", y = "Video score") +
    ggplot2::theme(legend.position = 'none',
                   axis.text.x = element_blank(),
                   axis.title.x = element_blank(),
                   axis.ticks.x = element_blank())


```

### Puzzle score

```{r}

# note we can use summary here instead of table because it is a continuous variable (not discrete)
summary(ice_cream$puzzle)

# histogram
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = puzzle) +
    ggplot2::geom_histogram(bins = 20, fill = "aquamarine4") +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "Puzzle score", y = "Count")

# redo histogram with less bins!
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = puzzle) +
    ggplot2::geom_histogram(bins = 10, fill = "aquamarine4") +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "Puzzle score", y = "Count")

# measures of spread
var(ice_cream$puzzle)
sd(ice_cream$puzzle)

# histogram with mean and sd plotted on top
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = puzzle) +
    ggplot2::geom_histogram(bins = 10, fill = "aquamarine4") +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "Video score", y = "Count") +
    ggplot2::geom_vline(xintercept = mean(ice_cream$puzzle), color = "blue", size = 2) +
    ggplot2::geom_vline(xintercept = c(mean(ice_cream$puzzle) + sd(ice_cream$puzzle),
                                       mean(ice_cream$puzzle) - sd(ice_cream$puzzle)), color = "black", size = 1.5, linetype = 2)

# box plot
ice_cream %>%
    ggplot2::ggplot(.) +
    ggplot2::aes(x = "1", y = puzzle) +
    ggplot2::geom_jitter(width = 0.2) +
    ggplot2::geom_boxplot(alpha = 0.5) +
    ggplot2::theme_bw(24) +
    ggplot2::labs(x = "", y = "Puzzle score") +
    ggplot2::theme(legend.position = 'none',
                   axis.text.x = element_blank(),
                   axis.title.x = element_blank(),
                   axis.ticks.x = element_blank())


```