---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 04

## Binomial random distributions

Which of the variables in our dataset could be classified as 'binomial'?

* binary
* independent
* fixed n
* same p

```{r}



```

If you randomly chose 5 individuals , what is the probability that that all five would be female?

```{r}

# calculate manually





# using dbinom



```

If you randomly chose 5 individuals , what is the probability that that only two would be female?

```{r}

# using dbinom(x, size, prob)



# using equation





```

If you randomly chose 5 individuals , what is the probability that that **at least two** would be female?

```{r}







```

## Normal distribution

Let's approximate the video scores as a normal distribution with a mean of `r mean(ice_cream$video)` and standard deviation `r sd(ice_cream$video)`. What is the probability that an individual scored lower than 60?

What is the probability that an individual scores between 40 and 57?

What is the probability than an individual scores over 50?

```{r}

# we can use pnorm with our mean and sd!



# pr(57) - pr(40) to get the range between values




# probability of scoring greater than 50 = 1-pr(50)



```

What is the 85th percentile of the video score data?

```{r}

# we have a probability, we need a quantile.



# check with pnorm



```

