---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```

# Lecture 04

## Binomial random distributions

Which of the variables in our dataset could be classified as 'binomial'?

* binary
* independent
* fixed n
* same p

```{r}
print("female!")
```

If you randomly chose 5 individuals , what is the probability that that all five would be female?

```{r}

# 109 females out of 200
table(ice_cream$female)

# probability that any one individual is female = 109/200
pr <- 109/200

# probability of multiple, independent events * number of possibilities
# pr(female) * pr(female) * pr(female) * pr(female) * pr(female) * choose(5,5)
pr**5 * choose(5,5)

# using dbinom(x, size, prob)
dbinom(5, 5, 109/200)

```

If you randomly chose 5 individuals , what is the probability that that only two would be female?

```{r}

# using dbinom(x, size, prob)
dbinom(2, 5, 109/200)

# using equation
# Pr(females)^(num females)*(Pr(males)^(num males))*combinatorial
(pr**2)*(1-pr)**3*(choose(5,2))

```

If you randomly chose 5 individuals , what is the probability that that **at least two** would be female?

```{r}

# 109 females out of 200
table(ice_cream$female)

# probability that any one individual is female = 109/200
pr <- 109/200

# using dbinom(x, size, prob)
# AT LEAST 2 = 2 OR 3 OR 4 OR 5
dbinom(2, 5, 109/200) + dbinom(3, 5, 109/200) + dbinom(4, 5, 109/200) + dbinom(5, 5, 109/200)

# alternatively, probability of NOT getting 1 or 0
1-(dbinom(1, 5, 109/200)+ dbinom(0, 5, 109/200))

# using pbinom
1 - pbinom(1,5,109/200)

```

## Normal distribution

Let's approximate the video scores as a normal distribution with a mean of `r mean(ice_cream$video)` and standard deviation `r sd(ice_cream$video)`. What is the probability that an individual scored lower than 60?

What is the probability that an individual scores between 40 and 57?

What is the probability than an individual scores over 50?

```{r}

# we can use pnorm with our mean and sd!
pnorm(60, mean = mean(ice_cream$video), sd = sd(ice_cream$video))

# pr(57) - pr(40) to get the range between values
pnorm(57, mean = mean(ice_cream$video), sd = sd(ice_cream$video))-pnorm(40, mean = mean(ice_cream$video), sd = sd(ice_cream$video))

# probability of scoring greater than 50 = 1-pr(50)
1-pnorm(50, mean = mean(ice_cream$video), sd = sd(ice_cream$video))

```

What is the 85th percentile of the video score data?

```{r}

# we have a probability, we need a quantile.
# use qnorm instead of pnorm!

qnorm(0.85, mean = mean(ice_cream$video), sd = sd(ice_cream$video))

# check with pnorm
pnorm(62.11, mean = mean(ice_cream$video), sd = sd(ice_cream$video))

```

