---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

# nice package/function to view pretty dataframe in R markdown
DT::datatable(ice_cream)

```

# Lecture 02

## Probability

What is the probability of choosing an individual whose favorite flavor is not vanilla?

```{r}

# 3 ways to solve this problem...



```

What is the probability that a student chosen at random is male?

```{r}

 



```

What is the probability of selecting an individual that is both male and enjoys chocolate?

* disjoint or non-disjoint?
* independent or dependent?
* AND or OR?

```{r}







```
