---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

# nice package/function to view pretty dataframe in R markdown
DT::datatable(ice_cream)

```

# Lecture 02

## Probability

What is the probability of choosing an individual whose favorite flavor is not vanilla?

```{r}

table(ice_cream$ice_cream)

# option 1: prob of not vanilla = 1 - pr(vanilla)!
1-(95/200)

# option 2: prob of not vanilla = pr(choc) + pr(straw)
(47/200) + (58/200)

# or freq(choc + straw) / total
(47+58)/200

```

What is the probability that a student chosen at random is male?

```{r}

# prob = relative frequency
table(ice_cream$female)

# pr(male) = # males / # students
91/200

```

What is the probability of selecting an individual that is both male and enjoys chocolate?

* disjoint or **non-disjoint**?
* **independent** or dependent? - as far as we know?
* **AND** or OR?

```{r}

# non-disjoint. (not mutually exclusive)
# as far as we know, they are independent but they might not be...
# and (multiply)

# pr(choc + male) = pr(choc) * pr(male | choc)
# pr(male | choc) = pr(male) if these events are independent.

# chocolate is 2 - 47/200 people like chocolate
table(ice_cream$ice_cream)

# pr(choc + male) = pr(choc) * pr(male) 
(47/200)*(91/200)

# look it up in the table... how many male choc lovers do we have? is this independent?
ice_cream %>%
    dplyr::mutate(ice_cream = dplyr::recode(ice_cream, "1" = "Vanilla", 
                                        "2" = "Chocolate",
                                        "3" = "Strawberry")) %>%
    dplyr::mutate(female = dplyr::recode(female, "0" = "Male", "1" = "Female")) %>%
    dplyr::group_by(ice_cream, female) %>%
    dplyr::summarize(n = n()) %>%
    DT::datatable()

#15 male chocolate lovers out of 47 total chocolate lovers
15/47

# not independent!!! pr(male) != pr(male | choc) --> we will discuss this more next time :)

```
