---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```


# Lecture 05

## Sampling distribution

Take a random sample of 5 individuals' video scores and calculate the sample mean and standard deviation. Does it match to your expectations?

```{r}

# take a sample
video_sample <- sample(ice_cream$video, size = 5)

# mean of population = 52.405 = mean sample


# sd of population = 10.7 = sd of sample


```

Now, take 100 samples each of 5 individuals video scores. Calculate the mean and standard deviation of the sampling distribution. Does it match your expectations?

```{r}
# create a variable to hold all the sample means

# take 10 samples, let "sam" be the variable for sample 1 or 2 or 3 etc.
# try a for loop!?



# look at means


# what is the mean of means?


# mean of population = 52.405 = mean sample

# what is the sd of means?


# sd of population = 10.7, sd of sample = sd/sqrt(n)



# plot the distribution of means?




```

Now try with 100 samples of 20 individuals video scores!

```{r}
# create a variable to hold all the sample means


# take 10 samples, let "sam" be the variable for sample 1 or 2 or 3 etc.





# look at means


# what is the mean of means?


# mean of population = 52.405 = mean sample

# what is the sd of means?


# sd of population = 10.7, sd of sample = sd/sqrt(n)


# plot the distribution of means?


# tighter distribution! Take care - the x axis is different scale so it looks less normal but its actually not! try to plot on same axes

```


If you take any given sample of 20 individuals' video scores, what is the probability that the average video score will be greater than 58?

```{r}

# use pnorm!



```
