---
title: "ice_cream_example"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(DT)
```

# Read in the data

* id: Identity of student
* female: Gender (0: male, 1: female)
* ice_cream: Favorite flavor (0: vanilla, 1: chocolate, 2: strawberry)
* video: score on video game
* puzzle: score on puzzle

```{r}

ice_cream <- read.csv("https://raw.githubusercontent.com/katiesevans/IGP_biostatistics/main/data/ice_cream.csv")

DT::datatable(ice_cream)

```


# Lecture 05

## Sampling distribution

Take a random sample of 5 individuals' video scores and calculate the sample mean and standard deviation. Does it match to your expectations?

```{r}

video_sample <- sample(ice_cream$video, size = 5)

# mean of population = 52.405 = mean sample
mean(ice_cream$video)
mean(video_sample) # yes!

# sd of population = 10.7 = sd of sample
sd(ice_cream$video)


```

Now, take 100 samples each of 5 individuals video scores. Calculate the mean and standard deviation of the sampling distribution. Does it match your expectations?

```{r}
# create a variable to hold all the sample means
means <- c()

# take 10 samples, let "sam" be the variable for sample 1 or 2 or 3 etc.
for(sam in 1:100) {
    # do a sample
    x <- sample(ice_cream$video, 5)
    # take the mean - add it to the means variable
    means <- c(means, mean(x))
}

# look at means
means

# what is the mean of means?
mean(means)

# mean of population = 52.405 = mean sample

# what is the sd of means?
sd(means)

# sd of population = 10.7, sd of sample = sd/sqrt(n)
10.7/sqrt(5) # this = 4.78 all good!

# plot the distribution of means?
hist(means)



```

Now try with 100 samples of 20 individuals video scores!

```{r}
# create a variable to hold all the sample means
means <- c()

# take 10 samples, let "sam" be the variable for sample 1 or 2 or 3 etc.
for(sam in 1:100) {
    # do a sample
    x <- sample(ice_cream$video, 20)
    # take the mean - add it to the means variable
    means <- c(means, mean(x))
}

# look at means
means

# what is the mean of means?
mean(means)

# mean of population = 52.405 = mean sample

# what is the sd of means?
sd(means)

# sd of population = 10.7, sd of sample = sd/sqrt(n)
10.7/sqrt(20) # this = 2.39 all good!

# plot the distribution of means?
hist(means)

# tighter distribution! Take care - the x axis is different scale so it looks less normal but its actually not! try to plot on same axes

```


If you take any given sample of 20 individuals' video scores, what is the probability that the average video score will be greater than 58?

```{r}

# use pnorm!
1-pnorm(58, mean = mean(ice_cream$video), sd = sd(ice_cream$video)/sqrt(20))

```

# Bonus: Plot a QQ plot and test the normality of the video scores and the puzzle scores.

```{r}

# plot qq plot for video
qqnorm(ice_cream$video)
qqline(ice_cream$video)

# shapiro-wilks test for non normality
shapiro.test(ice_cream$video)
# p < 0.05 - moderate evidence of non-normality

# plot qq plot for puzzle
qqnorm(ice_cream$puzzle)
qqline(ice_cream$puzzle)

# shapiro-wilks test for non normality
shapiro.test(ice_cream$puzzle)
# p < 0.001 - very strong evidence of non-normality


```
